Inference for Trans-dimensional Bayesian Models with Diffusive Nested Sampling

### 2014-12-05 Documents/Paper/ms.tex
New sinewave dataset

\subsection{Simulated Dataset}
To demonstrate the techniques,
we simulated some data based on the assumption $N=2$, i.e. the true signal was
composed of two sinusoids.
The simulated data is shown in Figure~\ref{fig:sinewave_data}, and shows a
large, low period oscillation with a smaller, much faster oscillation
superimposed. The noise level is such that the fast oscillation is difficult
to detect. The true values of the parameters were
$N=2$, $\mathbf{A} = \{1, 0.3\}$,
$\mathbf{T}=\{30, 2\}$, and $\boldsymbol{\phi} = \{0, 1\}$. The signal was
observed at $n=1001$ points equally spaced between $t=0$ and $t=100$.
The true form of the signal is:
\begin{eqnarray}
y(t) = \sin\left(\frac{2\pi t}{30}\right) +
0.3 \sin\left(\frac{2\pi t}{2} + 1\right)
\end{eqnarray}
and the noise standard deviation was $\sigma = 1$.

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{sinewave_data.pdf}
\caption{The simulated data for the sinusoidal example. The solid line shows
the true signal and the points are the measurements, simulated based on a
noise standard deviation of $\sigma = 1$.
\label{fig:sinewave_data}}
\end{center}
\end{figure}

@@

\subsection{Results}
We ran DNS on the simulated data shown in Figure~\ref{fig:sinewave_data}.
The shape of the likelihood function with respect to the prior mass (the
usual likelihood curve associated with Nested Sampling) is shown in
Figure~\ref{fig:sinewaves_likelihood}. Two phase transitions can be seen,
where this curve is concave up. The
first phase transition, at $\log(X) \approx -10$ nats,
separates ``noise-only'' models (where the
entire dataset is accounted for by the noise term) from ``one-sinusoid'' models
where $N=1$. At $\log(X) \approx -28$ nats, a second phase transition separates
$N=1$ models from $N=2$ models. This phase transition causes a double-peaked
feature in the posterior weights (lower panel of
Figure~\ref{fig:sinewaves_likelihood}).

>>>>

\subsection{Simulated Dataset}
To demonstrate the techniques,
we simulated some data based on the assumption $N=2$, i.e. the true signal was
composed of two sinusoids.
The simulated data is shown in Figure~\ref{fig:sinewave_data}, and shows a
large, low period oscillation with a smaller, much faster oscillation
superimposed. The noise level is such that the fast oscillation is difficult
to detect. The true values of the parameters were
$N=2$, $\mathbf{A} = \{7, 0.155\}$,
$\mathbf{T}=\{30, 2\}$, and $\boldsymbol{\phi} = \{0, 1\}$. The signal was
observed at $n=1001$ points equally spaced between $t=0$ and $t=100$.
The true form of the signal is:
\begin{eqnarray}
y(t) = \sin\left(\frac{2\pi t}{30}\right) +
0.3 \sin\left(\frac{2\pi t}{2} + 1\right)
\end{eqnarray}
and the noise standard deviation was $\sigma = 0.5$.

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{sinewave_data.pdf}
\caption{The simulated data for the sinusoidal example. The solid line shows
the true signal and the points are the measurements, simulated based on a
noise standard deviation of $\sigma = 0.5$.
\label{fig:sinewave_data}}
\end{center}
\end{figure}

@@

\subsection{Results}
We ran DNS on the simulated data shown in Figure~\ref{fig:sinewave_data}.
The shape of the likelihood function with respect to the prior mass (the
usual likelihood curve associated with Nested Sampling) is shown in
Figure~\ref{fig:sinewaves_likelihood}. Two phase transitions can be seen,
where this curve is concave up. The
first phase transition, at $\log(X) \approx -10$ nats,
separates ``noise-only'' models (where the
entire dataset is accounted for by the noise term) from ``one-sinusoid'' models
where $N=1$. At $\log(X) \approx -35$ nats, a second phase transition separates
$N=1$ models from $N=2$ models. This phase transition causes a double-peaked
feature in the posterior weights (lower panel of
Figure~\ref{fig:sinewaves_likelihood}).
